Mapping Cancer: the Potential of Cartograms and Alternative Map Displays

### 2020-12-08 ACE-19-31.Rmd
Clean repo for publication

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
  fig_path = "figs/",
  out.width = "\\textwidth")
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1. A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atlases are described in Table 2.", fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

@@

```{r ggcartograms, fig.cap = "Figure 2. Common alternatives to maps, showing the same information for the United States of America. The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016: (a) contiguous cartogram distorted each state's shape by the population of the state in 2015, (b) non-contiguous cartogram preserves the shape of the cartogram, but the size now reflects the population of the state in 2015 (c) Dorling cartogram are non-contiguous, circles are used to represent each state and the population of the state determines the size in 2015 (d) hexagon tilemap (non-contiguous) uses a hexagon of equal size for each state, and colours the state by the average age-adjusted rate of incidence. ", message=FALSE, warning=FALSE}
knitr::include_graphics("figures/usa_grid.png")
```

@@

```{r ggtilemap, fig.cap = "Figure 3. Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA. In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence. The geo-faceted map shows bar charts laid out in a grid approximating the spatial location of the state. The maps show age-adjusted rates for males and females. This display allows the presentation of multiple variables for each geographic area.", out.width="100%"}
knitr::include_graphics("figures/gggrids.png")
```

@@

```{r auscartograms, fig.cap = "Figure 4. Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/auscartograms.png")
```

@@

```{r interacting, fig.cap = "Figure 5. Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/interacting.png")
```

@@

```{r animating, fig.cap = "Figure 6. Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in Map of Cancer Mortality Rates in Spain, b. A highly responsive magnifying glass on a map of Breast Cancer Mortality in Canada.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/animating.png")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
  out.width = "\\textwidth")
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1. A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atlases are described in Table 2.", fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figs/F1 choropleth_grid.png")
```

@@

```{r ggcartograms, fig.cap = "Figure 2. Common alternatives to maps, showing the same information for the United States of America. The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016: (a) contiguous cartogram distorted each state's shape by the population of the state in 2015, (b) non-contiguous cartogram preserves the shape of the cartogram, but the size now reflects the population of the state in 2015 (c) Dorling cartogram are non-contiguous, circles are used to represent each state and the population of the state determines the size in 2015 (d) hexagon tilemap (non-contiguous) uses a hexagon of equal size for each state, and colours the state by the average age-adjusted rate of incidence. ", message=FALSE, warning=FALSE}
knitr::include_graphics("figs/F2 usa_grid.png")
```

@@

```{r ggtilemap, fig.cap = "Figure 3. Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA. In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence. The geo-faceted map shows bar charts laid out in a grid approximating the spatial location of the state. The maps show age-adjusted rates for males and females. This display allows the presentation of multiple variables for each geographic area.", out.width="100%"}
knitr::include_graphics("figs/F3 grids.png")
```

@@

```{r auscartograms, fig.cap = "Figure 4. Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figs/F4 auscartograms.png")
```

@@

```{r interacting, fig.cap = "Figure 5. Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figs/F5 interacting.png")
```

@@

```{r animating, fig.cap = "Figure 6. Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in Map of Cancer Mortality Rates in Spain, b. A highly responsive magnifying glass on a map of Breast Cancer Mortality in Canada.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figs/F6 animating.png")
```
